---
title: 'Statistical Inference Course Project - Part 2: Simulations'
author: "Alejandro Osorio"
date: "March 28, 2018"
output:
  pdf_document: default
  html_notebook: default
---

We're going to analyze the ToothGrowth data in the R datasets package.

### General Parameters
```{r, echo=FALSE}
# Parameters
library(tidyverse)
```

```{r}
data("ToothGrowth")
```

### Basic exploratory data analysis

Starting with a quick view at the table structure:
```{r}
head(ToothGrowth)
```

So it consists of 3 variables: Length, Supplement and Dose.

Now a quick glance at the table's size:
```{r}
dim(ToothGrowth)
```

So it's a small sample, therefore it'll require Student's test.

Just in case, let's check it there are any NAs in there:
```{r}
table(is.na(ToothGrowth))
```

So there are no NAs.  Now a quick glance at the variable's attributes:
```{r}
str(ToothGrowth)
```


It shows length as a numeric variable, the supplements as 2 types, stored as a factor ("OJ" and "VC"), and the dose as a repeating number.  Let's check how supplements are related to which dosages then:
```{r}
with(ToothGrowth,table(dose, supp))
```

As we can see, the sample consists of 3 dosages, for which there are 10 records of each supplement.

Additionally, a summary of the data to check any further information of interest:
```{r}
summary(ToothGrowth)
```

It shows a clear difference between the minimum and maximum tooth length, suggesting there's a good reason for some hypotheses to check.

Finally, a visual analysis of the data, using boxplots by supplement and dose, to check for evident patterns with which to work later.
```{r}
ggplot(ToothGrowth, aes(x = supp, y = len)) +
        geom_boxplot(aes(fill = supp), alpha = 0.5) +
        facet_wrap(~ dose, nrow = 1)
```

It suggests a clear relation between both supplement and dosage, with length.

The questions that arise from the exploratory analysis are, then:

1.- Can it be infered that, with higher dosages, longer teeth will be attained?

2.- If point 1 results to be true, Which supplement is more effective (obtains longer teeth)?

### Hypotheses testing

Both questions are tackled through hypothesis testing, based on the following assumptions:

a) Variables are iid
b) Samples are not paired from one dosage to the next (different subjects were monitored for each trial, with no relation whatsoever between them).
c) Underlying distributions for each sample, along dosages and supplements, have different variances.

#### Question 1: Higher dosages imply longer teeth?

The null hypothesis, for this case, is:  Teeth growth is not affected by dosage.

The alternative hypothesis is that, the higher the dosage, the longer teeth growth obtained.

In order to cover the whole spectrum of dosages (0.5, 1 and 2), two hypothesis testings are carried on:  a) one for the dosage interval of 0.5 - 1. b) the second one for the interval of 1 - 2, as follows:

##### Hypothesis testing for range 0.5 - 1

Selected samples are those associated to both dosages of 0.5 and dosages of 1, as follows:
```{r}
dose05 <- filter(ToothGrowth, ToothGrowth$dose == 0.5) %>%
        .$len
dose1 <- filter(ToothGrowth, ToothGrowth$dose == 1) %>%
        .$len
```

The one sided Student's test, then, based on the previous assumptions (included by default in the t.test) plus the filtered samples, goes as follows:
```{r}
t.test(dose1, dose05)
```

Which shows a greater teeth growth for dose1 (mean = 19.73) over dose05 (mean = 10.61), with a low enough p-value for the alternative hypothesis to replace the null one.

##### Hypothesis testing for range 1 - 2

Based on the same previous logic, the extra required sample, for dosage of 2 is obtained as follows:
```{r}
dose2 <- filter(ToothGrowth, ToothGrowth$dose == 2) %>%
        .$len
```

The one sided Student's test, then, goes as follows:
```{r}
t.test(dose2, dose1)
```

Which shows a greater teeth growth for dose2 (mean = 26.10) over dose1 (mean = 19.74), with a low enough p-value for the alternative hypothesis to, again, replace the null one.

#### Question 2: Which supplement is more effective (obtains longer teeth)?

The null hypothesis, for this case, is:  Teeth growth is the same for both supplements.

The alternative hypothesis is that, OJ supplement is more effective than VC.

In this case, just one hypothesis testing is carried on, with the samples being those associated to OJ and VC supplements respectively, as follows:

```{r}
suppoj <- filter(ToothGrowth, ToothGrowth$supp == "OJ") %>%
        .$len
suppvc <- filter(ToothGrowth, ToothGrowth$supp == "VC") %>%
        .$len
```

The one sided Student's test, then, based on the previous assumptions (included by default in the t.test) plus the filtered samples, goes as follows:
```{r}
t.test(suppoj, suppvc)
```

Which shows higher effectiveness of supplement OJ (mean = 20.66), over supplement VC (mean = 16.96) with a low enough p-value for the alternative hypothesis to replace the null one.

### Final Conclusions

Based on the previous preliminary data analysis, plus hypotheses testing, it can be concluded that:

1.- The higher the dosage, the longer the teeth growth obtained.

2.- Supplement OJ is more effective than supplement VC.
